Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-12 article/article.Rmd
Commit 2024-06-12 10:14

The distribution of the number of CE events as a function of OIR is shown in
Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently, there is an inverse
relation between the frequency of CE events  and OIR with an exponential-like
decline. This is expected, as the stronger the CE events become, the rarer are the
conditions favoring the occurrence of CE events. For the majority (over $60\%$) of
the CE events the OIR is below the long term average of
$`r round(mean(ST_E_yearly$GLB_diff.mean), 1)`\,\text{W}/\text{m}^2$,
while about $10\%$ of the events correspond to OIR larger than
$100\,\text{W}/\text{m}^2$ and up to highest value of 300 W/m^2^.
This distribution is indicative of the magnitude and the probability of the expected
CE
events over Thessaloniki. Similar distribution of CE events, occurrences have been reported
by @Vamvakas2020, where the magnitude of OIR were
higher due to the location of the city of Patras, $2.5^\circ$ closer to the equator.

```{r ovir-distribution, echo=FALSE, fig.cap="Distribution of CE over irradiance magnitude."}
knitr::include_graphics("../images/P-relative-distribution-diff-1.png")
```

<!-- whisky plots relative to GLB -->

<!-- set of yearly plots -->

## Groups of cloud enhancement

In order to study the total duration of the CE events, we grouped the one
minute CE events to continuous CE groups (CEG). Thus, a CEG consists of one or more
successive CE cases, and can represent cloud enhancement conditions spanning a
duration of multiple minutes.  We have identified `r ST_G0[, .N]` groups of CE in the
whole period of study, where the
group with the longest duration lasted
`r ST_G0[, max(GLB_diff.N)]` minutes on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can conclude, that
although we detected some CEG with durations longer than an hour, about $80\,\%$ of
the CEG have a duration of less than 5 minutes.

>>>>

The distribution of the number of CE events as a function of OIR is shown in
Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently, there is an inverse
relation between the frequency of CE events  and OIR with an exponential-like
decline. This is expected, as the stronger the CE events become, the rarer are the
conditions favoring the occurrence of CE events. For the majority (over $62\%$) of
the CE events the OIR is below the long term average of
$`r round(mean(ST_E_yearly$GLB_diff.mean), 1)`\,\text{W}/\text{m}^2$,
while about $8.1\%$ of the events correspond to OIR larger than
$100\,\text{W}/\text{m}^2$ and up to highest value of $412.4\,\text{W}/\text{m}^2$.
This distribution is indicative of the magnitude and the probability of the expected
CE events over Thessaloniki.  Similar distribution of CE events, albeit with larger
OIR values, has been reported by @Vamvakas2020, fot the city of Patras.This site is
located $2.5^\circ$ closer to the equator.  south of Thessaloniki and is exposed to
air masses coming mainly from the eastern Mediterranean resulting in different cloud
patterns, that may affect the characteristics and magnitude of the CE events.

```{r ovir-distribution, echo=FALSE, fig.cap="Relative frequency distribution of CE events as a function of OIR."}
knitr::include_graphics("../images/P-relative-distribution-diff-1.png")
```

## Duration of cloud enhancement events

The duration of the CE events is variable and can last for several minutes or even
more than an hour. In order to study the characteristics of these consecutive events,
we grouped them into bins of increasing duration in steps of one minute.

We have identified `r ST_G0[, .N]` groups of CE in the
whole period of study, where the
group with the longest duration lasted
`r ST_G0[, max(GLB_diff.N)]` minutes on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can conclude, that
although we detected some CEG with durations longer than an hour, about $80\,\%$ of
the CEG have a duration of less than 5 minutes.
